---
title: "Data-preparation"
output: html_document
date: "2023-07-07"
---

## Libraries
```{r, message=FALSE}
# increase Java memory
options(java.parameters = "-Xmx6G")
#cancensus API key,
options(cancensus.api_key = "XX")

library(disk.frame)
library(cancensus)
library(sf)
library(ggplot2)
library(spatstat)
library(leaflet)
library(r5r)
library(dplyr)
```

# Origins (DA, their centroids, and census data)
Pulling data from {cancensus} for the Hamilton Census Subdivision (CSD) and various census variables at the level of the Dissemination Area (DA). We can see that the Hamilton CSD has a population of 569,353 and is identified by the CSD region id of "3525005". It is part of the CMA with ID 35537.
```{r}
list_census_regions('CA21') %>% 
  filter(level == "CSD", name %in% c("Hamilton"))
```

```{r}
list_census_regions('CA21') %>% 
  filter(level == "CSD", name %in% c("Hamilton"))
```

```{r, eval=FALSE}
HAM_census_21 <- get_census(dataset='CA21', 
                            regions=list(CSD="3525005"),
                            vectors=c("v_CA21_11", #total - aged 0-14
                                      "v_CA21_12", #males - aged 0-14
                                      "v_CA21_13", #females - aged 0-14
                                      "v_CA21_68", #total - aged 15-64
                                      "v_CA21_69", #males - aged 15-64
                                      "v_CA21_70", #females - aged 15-64
                                      "v_CA21_251", #total - aged 65 and over
                                      "v_CA21_252", #males - aged 65 and over
                                      "v_CA21_253", #females - aged 65 and over
                                      "v_CA21_1085", #total - prevalence of LICO-AT (%)
                                      "v_CA21_1086", #male - prevalence of LICO-AT (%)
                                      "v_CA21_1087",#female - prevalence of LICO-AT (%)
                                      "v_CA21_7647", #total - walked to work
                                      "v_CA21_7635", #total - drove or was driven to work
                                      "v_CA21_7644", #total - used transit to get to work
                                      "v_CA21_7650", #total - biked to work
                                      "v_CA21_7653" #total - used some other mode to get to work
                                      ), 
                                      level='DA', use_cache = FALSE, geo_format = 'sf', quiet = TRUE)

#NOTE: LICO-AF = "The Low‑income cut‑offs, after tax refer to income thresholds, defined using 1992 expenditure data, below which economic families or persons not in economic families would likely have devoted a larger share of their after‑tax income than average to the necessities of food, shelter and clothing. More specifically, the thresholds represented income levels at which these families or persons were expected to spend 20 percentage points or more of their after‑tax income than average on food, shelter and clothing. " 

#NOTE: the 'mode choice' is the count of the TOTAL population aged 15 years and over who is employed in the labour force.
```

Check:
```{r}
# Mapping DAs
leaflet() |>
  addTiles() |>
  addPolygons(data= HAM_census_21)
```

Save DA census data (sf object, polygons) :
```{r, eval=FALSE}
save(HAM_census_21, file="data-intermediate/HAM_census_21.rda")
```

Let's also pull the 2016 data:
```{r, eval=FALSE}
HAM_census_16 <- get_census(dataset='CA16', 
                            regions=list(CSD="3525005"),
                            vectors=c("v_CA16_1", #total pop
                                      "v_CA16_2", #total pop - males
                                      "v_CA16_3", #total pop - females
                                      "v_CA16_2570", #total - prevalence of LICO-AT (%)
                                      "v_CA16_2571", #male - prevalence of LICO-AT (%)
                                      "v_CA16_2572", #female - prevalence of LICO-AT (%)
                                      "v_CA16_5810", #total - used some other mode to get to work
                                      "v_CA16_5804", #total - walked to work
                                      "v_CA16_5795", #total - car as driver 
                                      "v_CA16_5798", #total - car as passenger
                                      "v_CA16_5801", #total - used transit to get to work
                                      "v_CA16_5807" #total - biked to work
                                      ), 
                                      level='DA', use_cache = FALSE, geo_format = 'sf', quiet = TRUE)

#NOTE: LICO-AF = "The Low‑income cut‑offs, after tax refer to income thresholds, defined using 1992 expenditure data, below which economic families or persons not in economic families would likely have devoted a larger share of their after‑tax income than average to the necessities of food, shelter and clothing. More specifically, the thresholds represented income levels at which these families or persons were expected to spend 20 percentage points or more of their after‑tax income than average on food, shelter and clothing. " 

#NOTE: the 'mode choice' is the count of the TOTAL population aged 15 years and over who is employed in the labour force.


#Check: Mapping DAs
leaflet() |>
  addTiles() |>
  addPolygons(data= HAM_census_16)
```

Save DA census data (sf object, polygons) :
```{r, eval=FALSE}
save(HAM_census_16, file="data-intermediate/HAM_census_16.rda")
```


# Care destinations
Load the *care destination* database that has been created and reformat it as a sf objectL
```{r}
care_dest <- read.csv("data-raw/FINAL_Care_Destinations_2023.csv")

care_dest <- st_as_sf(care_dest, coords = c("LONGITUDE", "LATITUDE"))
```

Check:
```{r}
leaflet() |>
  addTiles() |>
  addMarkers(data= care_dest, popup = ~as.character(Name))
```

Save care destinations (sf object, points):
```{r}
save(care_dest, file="data-intermediate/care_dest.rda")
```

# Travel time calculations - from DA centroids to care destinations, all modes.

Formatting origin (centroids of DAs (from_ids)) and destination (centroids of care destinations) points for the travel time calcs using r5r:
```{r calc centroids}
orig <- st_centroid(HAM_census_21) #taking the geometric centroids of each DA. This will be the 'origin'

orig$lon <- st_coordinates(orig)[,1] 
orig$lat <- st_coordinates(orig)[,2]

orig <- orig %>% 
  st_drop_geometry() %>%
  transmute(id = GeoUID,
            lon,
            lat)

dest <- care_dest

dest$lon <- st_coordinates(dest)[,1]
dest$lat <- st_coordinates(dest)[,2]

dest <- dest %>% 
  st_drop_geometry() %>%
  transmute(ID = as.character(ID),
            lon,
            lat) %>%
  rename("id" = "ID")
```

```{r set up r5 path, include=FALSE}
# the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.html . See the direction given in the r5r installation here (https://ipeagit.github.io/r5r/articles/r5r.html)
dir.create("./data-raw/tt") #create a folder, if it already exists - this function does nothing

r5_path <- file.path("./data-raw/tt")
list.files(r5_path)
```

Download OSM Network:
```{r download OSM street network, eval=FALSE}
# Specify the URL of the OSM file for Ontario - it takes a few minutes to download (~56MB)
download.file(url = "http://download.geofabrik.de/north-america/canada/ontario-latest.osm.pbf",
              destfile= file.path(r5_path, "Hamilton.osm.pbf"), mode = "wb")
```
```{r}
#clip to the the hamilton city limits using OSM Convert .exe (). (NOTE: x = longitude)
osmdata::getbb("Hamilton")
```

Download GTFS data:
```{r download GTFS data, eval=FALSE}
download.file(url = "https://transitfeeds.com/p/hamilton-street-railway/31/latest/download",
              destfile= file.path(r5_path, "HSR_transit.zip"), mode = "wb")
```

Set-up R5 Routing:
```{r build graph, eval = FALSE}
#takesa few minutes
r5_HAM <- setup_r5(data_path = r5_path, verbose = TRUE)
```

Create map of transportation network for paper:
```{r}
#boundaries for mapping
ham_comms <- st_read(paste0(here::here(),"/data-raw/Boundaries/Community_Boundaries.shp")) |> mutate(COMMUNITY_ = ifelse(COMMUNITY_ == "Hamilton", "Hamilton-Central", COMMUNITY_))  |> st_transform(4326)
ham_bound <- st_read(paste0(here::here(),"/data-raw/Boundaries/City_Boundary.shp")) |> st_transform(4326)

hydro_p_LakeOntario <- st_read(paste0(here::here(),"/data-raw/Boundaries/hydro_p_LakeOntario.shp")) |> st_transform(crs=4326)
ham_bay <- st_read(paste0(here::here(),"/data-raw/Boundaries/Waterbodies.shp")) |> st_transform(crs=4326) |> filter(FEATURE_TY == "Lake")
```

```{r}
# extract OSM network
street_net <- street_network_to_sf(r5_HAM)
# extract public transport network
transit_net <- r5r::transit_network_to_sf(r5_HAM)

# simplify
transit_stops <- transit_net$stops 
car_net <- street_net$edges |> filter(car == TRUE)
bike12_or_walk_net <- street_net$edges |> filter(bicycle_lts == 1 | bicycle_lts == 2 )

transit_stops <- st_intersection(transit_stops, ham_bound)
car_net <- st_intersection(car_net, ham_bound)
bike12_or_walk_net <- st_intersection(bike12_or_walk_net, ham_bound)
car_net_filtered <- car_net |> filter(street_class == "PRIMARY" | street_class == "MOTORWAY" | street_class == "SECONDARY")
```

```{r}
transit_stops_coords <- data.frame(lat = st_coordinates(transit_stops)[,1],
                                   lon = st_coordinates(transit_stops)[,2])

#osmdata::getbb("Hamilton") #to get the extent of hamilton to set thexlim and ylim

```


```{r, eval=FALSE}
descriptive_ham_plot <- ggplot() +
  geom_sf(data = ham_bound, fill="white")+
  geom_sf(data = ham_bay, fill="skyblue", color = "skyblue")+
  geom_sf(data = hydro_p_LakeOntario, fill="skyblue", color = "skyblue")+
  geom_bin2d(data = transit_stops_coords,
                   aes(x = lat,
                       y = lon)) +
  geom_sf(data = car_net_filtered, color='gray30', linewidth=0.3, alpha=0.5) +
  geom_sf(data = bike12_or_walk_net,  color='gray80', linewidth=0.05, alpha=0.6) +
  geom_sf(data = ham_comms, fill=NA, color="#41AB5D", linewidth=0.6)+
  scale_fill_gradient(high = "#CB181D", low= "#FCBBA1", name="Number. of\nbus stops", breaks = c(1, 25, 50, 75, 100))+ 
  ggspatial::annotation_scale(location = "bl") +
  ggspatial::annotation_north_arrow(location = "tl", height = unit(1.0, "cm"), width = unit(1.0, "cm")) +
  xlim(c(-80.24856, -79.62211))+
  ylim(c(43.05051, 43.47068))+
  theme_void()+
  theme(panel.background = element_rect(fill = "grey"),
        legend.position = "inside",
        legend.justification.inside = c(1,1),
        legend.position.inside = c(1,1))

descriptive_ham_plot

ggsave(file = paste0(here::here(),"/Journal-manuscript/figures/Fig1-descriptive-boundaries.png"), dpi=300)
```



Calculate **car** travel times from centroids of DAs (from_ids) to destination points (to_ids). 60 min max travel time. 8am departure on a Tuesday. 30min +/- departure window. 
```{r car travel time, message=FALSE, eval=FALSE}
departure_datetime <- as.POSIXct("08-06-2023 08:00:00 EDT", # 12 GMT is 8am in Toronto
                                 format = "%d-%m-%Y %H:%M:%S")


start.time <- Sys.time()
ttm_care_car <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = orig,
                          destinations = dest,
                          mode = c("CAR"),
                          departure_datetime = departure_datetime,
                          time_window = 30,
                          percentiles = c(25,50,75),
                          max_trip_duration = 60,
                          verbose = TRUE)

end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_car <- as.data.frame(ttm_care_car)

save("ttm_care_car", file = "data-raw/tt/ttm_care_car.rda")
```

Calculate **transit** travel times from centroids of DAs (from_ids) to destination points (to_ids). 60 min max travel time, 15 min max walking limit. 8am departure on a Tuesday. 30min +/- departure window.
```{r transit travel time}
departure_datetime <- as.POSIXct("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")


start.time <- Sys.time()
ttm_care_transit <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = orig,
                          destinations = dest,
                          mode = c("TRANSIT"),
                          mode_egress = "WALK",
                          departure_datetime = departure_datetime,
                          time_window = 30,
                          percentiles = c(25,50,75),
                          max_trip_duration = 60,
                          max_walk_time = 15,
                          verbose = TRUE)

end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_transit <- as.data.frame(ttm_care_transit)

save("ttm_care_transit", file = "data-raw/tt/ttm_care_transit.rda")
```

Calculate **walk** travel times from centroids of DAs (from_ids) to destination points (to_ids). 60 min max travel time, 15 min max walking limit. 8am departure on a Tuesday. 30min +/- departure window.
```{r walk travel time}
departure_datetime <- as.POSIXct("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")


start.time <- Sys.time()
ttm_care_walk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = orig,
                          destinations = dest,
                          mode = c("WALK"),
                          departure_datetime = departure_datetime,
                          time_window = 30,
                          percentiles = c(25,50,75),
                          max_trip_duration = 60,
                          verbose = TRUE)

end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_walk <- as.data.frame(ttm_care_walk)

save("ttm_care_walk", file = "data-raw/tt/ttm_care_walk.rda")
```

Calculate **bike** travel times from centroids of DAs (from_ids) to destination points (to_ids). 60 min max travel time, 15 min max walking limit. 8am departure on a Tuesday. 30min +/- departure window.
```{r bike travel time}
departure_datetime <- as.POSIXct("08-06-2023 08:00:00", # 12 GMT is 8am in Toronto (EDT)
                                 format = "%d-%m-%Y %H:%M:%S")


start.time <- Sys.time()
ttm_care_bike <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = orig,
                          destinations = dest,
                          mode = c("BICYCLE"),
                          departure_datetime = departure_datetime,
                          time_window = 30,
                          percentiles = c(25,50,75),
                          max_trip_duration = 60,
                          verbose = TRUE)

end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_bike <- as.data.frame(ttm_care_bike)

save("ttm_care_bike", file = "data-raw/tt/ttm_care_bike.rda")
```

Limitations of travel time calcs:
- does not consider traffic stress (i.e., car, bike, and walk for all percentiles are identical -- the time window does not matter and no traffic is assumed)
- for transit, the time window accounts for departures +/- 30 minutes. This accounts for transit availability variability but also does not accont for traffic stress

# Travel time calculations - from parcel points to destinations

Import new origins, this time they are parcels.
```{r}
load("data-raw/R_PARCELS_CENTS_2020.rda")

orig_sf <- R_PARCELS_CENTS_2020 |> st_as_sf()
rm(R_PARCELS_CENTS_2020,R_PARCELS_2020)

coords <- st_coordinates(orig_sf)
orig <- data.frame(id = orig_sf$ID,
                   lon = st_coordinates(orig_sf)[,1],
                   lat = st_coordinates(orig_sf)[,2])

rm(orig_sf)
```
```{r}
# due to size, the calculation of travel times needs to be batched. We split the origins into 4 distinct dataframes.

orig_1 <- orig[1:36903,]
orig_2 <- orig[36904:73807,]
orig_3 <- orig[73808:110710,]
orig_4 <- orig[110711:143893,]

#check, make sure sums to 143893
nrow(orig_1)+nrow(orig_2)+nrow(orig_3)+nrow(orig_4)
```

As an alternative, calculate **transit** travel times from parcel ids (from_ids) to destination points (to_ids). 30 min max travel time. 8am departure on a Tuesday. 30min +/- departure window. 


```{r Transit travel time 1}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_1)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_1,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = "TRANSIT",
                          mode_egress = "WALK",
                          max_walk_time = 15,
                          percentiles = c(25,50,75),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00",
                          format = "%d-%m-%Y %H:%M:%S")),
                          time_window = 60,
                          max_trip_duration = 30)
  
  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took ~60 minutes..."
ttm_care_TRANSIT_parcel1 <- as.data.frame(output_df)

save("ttm_care_TRANSIT_parcel1", file = "./data-intermediate/parcel_tts/ttm_care_TRANSIT_parcel1_1.rda")
rm("ttm_care_TRANSIT_parcel1_1")
```

```{r TRANSIT travel time 2}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_2)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_2,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = "TRANSIT",
                          mode_egress = "WALK",
                          max_walk_time = 15,
                          percentiles = c(25,50,75),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00",
                          format = "%d-%m-%Y %H:%M:%S")),
                          time_window = 60,
                          max_trip_duration = 30)
  
  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took ~60 minutes..."
ttm_care_TRANSIT_parcel2 <- as.data.frame(output_df)

save("ttm_care_TRANSIT_parcel2", file = "./data-intermediate/parcel_tts/ttm_care_TRANSIT_parcel2.rda")
rm("ttm_care_TRANSIT_parcel2")
```

```{r TRANSIT travel time 3}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_3)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_3,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = "TRANSIT",
                          mode_egress = "WALK",
                          max_walk_time = 15,
                          percentiles = c(25,50,75),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00",
                          format = "%d-%m-%Y %H:%M:%S")),
                          time_window = 60,
                          max_trip_duration = 30)
  
  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_TRANSIT_parcel3 <- as.data.frame(output_df)

save("ttm_care_TRANSIT_parcel3", file = "./data-intermediate/parcel_tts/ttm_care_TRANSIT_parcel3.rda")
rm("ttm_care_TRANSIT_parcel3")
```

```{r TRANSIT travel time 4}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_4)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_4,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = "TRANSIT",
                          mode_egress = "WALK",
                          max_walk_time = 15,
                          percentiles = c(25,50,75),
                          departure_datetime = as.POSIXct(strptime("08-06-2023 08:00:00",
                          format = "%d-%m-%Y %H:%M:%S")),
                          time_window = 60,
                          max_trip_duration = 30)
  
  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took ~60 minutes..."

ttm_care_TRANSIT_parcel4 <- as.data.frame(output_df)

save("ttm_care_TRANSIT_parcel4", file = "./data-intermediate/parcel_tts/ttm_care_TRANSIT_parcel4.rda")
rm("ttm_care_TRANSIT_parcel4")
```

Now let's calculate car travel times. Max. 30 min trip. from parcels.
```{r CAR travel time 1}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_1)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_1,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = "CAR",
                          percentiles = 50,
                          max_trip_duration = 30)
  
  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took ~60 minutes..."
ttm_care_CAR_parcel1 <- as.data.frame(output_df)

save("ttm_care_CAR_parcel1", file = "./data-intermediate/parcel_tts/ttm_care_CAR_parcel1_1.rda")
rm("ttm_care_CAR_parcel1_1")
```

```{r CAR travel time 2}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_2)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_2,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = "CAR",
                          percentiles = 50,
                          max_trip_duration = 30)
  
  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took ~60 minutes..."
ttm_care_CAR_parcel2 <- as.data.frame(output_df)

save("ttm_care_CAR_parcel2", file = "./data-intermediate/parcel_tts/ttm_care_CAR_parcel2.rda")
rm("ttm_care_CAR_parcel2")
```

```{r CAR travel time 3}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_3)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_3,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = "CAR",
                          percentiles = 50,
                          max_trip_duration = 30)
  
  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))

ttm_care_CAR_parcel3 <- as.data.frame(output_df)

save("ttm_care_CAR_parcel3", file = "./data-intermediate/parcel_tts/ttm_care_CAR_parcel3.rda")
rm("ttm_care_CAR_parcel3")
```

```{r CAR travel time 4}
# set up batching according to how many origin rows to process at one time
chunksize = 50 
num_chunks = ceiling(nrow(orig_4)/chunksize)

# create origin-destination pairs
origins_chunks <- as.disk.frame(orig_4,
                          outdir = "parcel-ttm/processing/orig",
                          nchunks = num_chunks,
                          overwrite = TRUE)

start.time <- Sys.time()
pb <- txtProgressBar(0, num_chunks, style = 3)

for (i in 1:num_chunks){
  Orig_chunk <- get_chunk(origins_chunks, i)
  ttm_chunk <- travel_time_matrix(r5r_core = r5_HAM,
                          origins = Orig_chunk,
                          destinations = dest,
                          mode = "CAR",
                          percentiles = 50,
                          max_trip_duration = 30)
  
  # export output as disk.frame
  ifelse(i == 1, output_df <- as.disk.frame(ttm_chunk,
                                            nchunks = 1,
                                            outdir = "parcel-ttm/processing/output_ttm",
                                            compress = 50,
                                            overwrite = TRUE),
         add_chunk(output_df, ttm_chunk, chunk_id = i))
  setTxtProgressBar(pb, i)
}
end.time <- Sys.time()
print(paste0("OD matrix calculation took ", round(difftime(end.time, start.time, units = "mins"), digits = 2), " minutes..."))
#"OD matrix calculation took ~60 minutes..."

ttm_care_CAR_parcel4 <- as.data.frame(output_df)

save("ttm_care_CAR_parcel4", file = "./data-intermediate/parcel_tts/ttm_care_CAR_parcel4.rda")
rm("ttm_care_CAR_parcel4")
```

