---
title: "Care-access-analysis"
output: html_document
date: "2023-07-09"
---
Libraries:
```{r libraries}
library(sf)
library(tmap)
library(dplyr)
library(biscale)
library(ggplot2)
```
NOTE: Used p25 travel times for transit mode.

# Load data
```{r load packages}
# DAs (origins)
load("data-intermediate/HAM_census_21.rda")

# Care destinations 
load("data-intermediate/care_dest.rda")

# travel time matrices
load("data-intermediate/ttm_care_car.rda")
load("data-intermediate/ttm_care_bus.rda")
load("data-intermediate/ttm_care_bike.rda")
load("data-intermediate/ttm_care_walk.rda")

# city boundaries
ham_bound <- read_sf("data-raw/Boundaries/City_Boundary.shp")
ham_comms <- read_sf("data-raw/Boundaries/Community_Boundaries.shp")
```

# Cumulative opportunity calculations
No. of care opportunities that can be reached within 30 mins or less:
```{r 30 min cumulative opps}
ttm_car_30 <- ttm_care_car %>% filter(travel_time_p50 <= 30) %>%
  group_by(from_id) %>% 
  summarise(cum_opps = n())

ttm_walk_30 <- ttm_care_walk %>% filter(travel_time_p50 <= 30) %>%
  group_by(from_id) %>% 
  summarise(cum_opps = n())

ttm_bike_30 <- ttm_care_bike %>% filter(travel_time_p50 <= 30) %>%
  group_by(from_id) %>% 
  summarise(cum_opps = n())

ttm_transit_30 <- ttm_care_transit %>% filter(travel_time_p25 <= 30) %>%
  group_by(from_id) %>% 
  summarise(cum_opps = n())
```

No. of care opportunities that can be reached within 15 mins or less:
```{r 15 min cumulative opps}
ttm_car_15 <- ttm_care_car %>% filter(travel_time_p50 <= 15) %>%
  group_by(from_id) %>% 
  summarise(cum_opps = n())

ttm_walk_15 <- ttm_care_walk %>% filter(travel_time_p50 <= 15) %>%
  group_by(from_id) %>% 
  summarise(cum_opps = n())

ttm_bike_15 <- ttm_care_bike %>% filter(travel_time_p50 <= 15) %>%
  group_by(from_id) %>% 
  summarise(cum_opps = n())

ttm_transit_15 <- ttm_care_transit %>% filter(travel_time_p25 <= 15) %>%
  group_by(from_id) %>% 
  summarise(cum_opps = n())
```

# Visualizing - 30 minute cumulative opportunities
```{r}
orig_full_sf <- HAM_census_21 %>% 
  select(GeoUID, 
         Population, 
         `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`) %>%  #add the census variable fields that you want to pull here
  rename("LICO_AT" = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`)
```

```{r}
HAM_access_30_1 <- orig_full_sf %>% 
  merge(ttm_car_30, by.x="GeoUID", by.y="from_id", all.x=TRUE) %>%
  mutate(mode = "a_car") 

HAM_access_30_2 <-orig_full_sf %>% 
  merge(ttm_transit_30, by.x="GeoUID", by.y="from_id",all.x=TRUE) %>%
  mutate(mode = "b_transit") 

HAM_access_30_3 <-orig_full_sf %>% 
  merge(ttm_bike_30, by.x="GeoUID", by.y="from_id", all.x=TRUE) %>%
  mutate(mode = "c_bike")

HAM_access_30_4 <-orig_full_sf %>% 
  merge(ttm_walk_30, by.x="GeoUID", by.y="from_id", all.x=TRUE) %>%
  mutate(mode = "d_walk")

HAM_access_30 <- rbind(HAM_access_30_1, HAM_access_30_2, HAM_access_30_3, HAM_access_30_4) %>%
  mutate(cum_opps = ifelse(is.na(cum_opps), 0, cum_opps))

remove(HAM_access_30_1,HAM_access_30_2,HAM_access_30_3,HAM_access_30_4)
```

```{r}
tm_shape(HAM_access_30) +
  tm_polygons("cum_opps",
              palette = "Purples",
              breaks = c(0,58,212, 665, 729, 2215),
              border.alpha = 0,
              title = "Cumulative care\nopportunities (30 min)") +
  tm_facets(by=c("mode"), nrow=2, showNA = FALSE)+
  tm_shape(ham_comms) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("right", "bottom"),breaks=c(0,1,2,4) )+
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_legend(legend.position = c("left", "center"))+
  tm_layout(panel.labels=c("Car",
                           "Transit",
                           "Bike",
                           "Walk"),
            panel.label.color = "black",
            panel.label.size=1.3,
            panel.label.fontface="bold")
```


# Visualizing - 15 minute cumulative opportunities
```{r}
HAM_access_15_1 <- orig_full_sf %>% 
  merge(ttm_car_15, by.x="GeoUID", by.y="from_id", all.x=TRUE) %>%
  mutate(mode = "a_car") 

HAM_access_15_2 <-orig_full_sf %>% 
  merge(ttm_transit_15, by.x="GeoUID", by.y="from_id",all.x=TRUE) %>%
  mutate(mode = "b_transit") 

HAM_access_15_3 <-orig_full_sf %>% 
  merge(ttm_bike_15, by.x="GeoUID", by.y="from_id", all.x=TRUE) %>%
  mutate(mode = "c_bike")

HAM_access_15_4 <-orig_full_sf %>% 
  merge(ttm_walk_15, by.x="GeoUID", by.y="from_id", all.x=TRUE) %>%
  mutate(mode = "d_walk")

HAM_access_15 <- rbind(HAM_access_15_1, HAM_access_15_2, HAM_access_15_3, HAM_access_15_4) %>%
  mutate(cum_opps = ifelse(is.na(cum_opps), 0, cum_opps))

remove(HAM_access_15_1,HAM_access_15_2,HAM_access_15_3,HAM_access_15_4)
```

```{r}
tm_shape(HAM_access_15) +
  tm_polygons("cum_opps",
              palette = "Oranges",
              breaks = c(0,11,41,234,329,1818),
              border.alpha = 0,
              title = "Cumulative care\nopportunities (15 min)",
              colorNA = "white", 
              colorNULL = "white" ) +
  tm_facets(by=c("mode"), nrow=2, showNA = FALSE)+
  tm_shape(ham_comms) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("right", "bottom"),breaks=c(0,1,2,4) )+
  tm_compass(position = c("left", "top"), size=1.0)+
  tm_legend(legend.position = c("left", "center"))+
  tm_layout(panel.labels=c("Car",
                           "Transit",
                           "Bike",
                           "Walk"),
            panel.label.color = "black",
            panel.label.size=1.3,
            panel.label.fontface="bold")
```
# Visualizing - % low income and population
The map data:
```{r}
data <- bi_class(HAM_census_21 %>% mutate(Pop_den = Population/`Shape Area`), 
                 x = Pop_den, 
                 y = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`,
                 style = "quantile", 
                 dim = 3)
```

The plot and legend displayed:
```{r}
map <- ggplot() +
  geom_sf(data = data, mapping = aes(fill = bi_class), 
          color = NA, size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "GrPink", dim = 3) +
  geom_sf(data = ham_comms, fill=NA) +
  bi_theme() +
  ggtitle("Population density and LICO %")

legend <- bi_legend(pal = "GrPink",
                    dim = 3,
                    xlab = "Pop D.",
                    ylab = "LICO % ",
                    size = 8)

#the final plot all put together
cowplot::ggdraw() +
  cowplot::draw_plot(map, 0, 0, 1, 1) +
  cowplot::draw_plot(legend, 0.6, .65, 0.2, 0.2)
```

