---
title: "Cancensus Centroid"
output: html_document
date: "2023-05-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("cancensus")
library(cancensus)
library(sf)
install.packages('geojsonsf')
library(geojsonsf)
options(cancensus.api_key = "CensusMapper_1d16ff8ad5e8a65ff028959182ac4940")
```
Hamilton CMA: "35537"

Income Vector:Income statistics in 2020 for the population aged 15 years and over in private households: "v_CA21_599"
Education Vector: Total - Highest certificate, diploma or degree for the population aged 15 years and over in private households - "v_CA21_5817"
Gender Vector:
Age Vector: Total Age - "v_CA21_8"
Ethnicity Vector: Total - Ethnic or cultural origin for the population in private households : "v_CA21_4917"]
```{r}
census_data <- get_census(dataset='CA21', regions=list(CMA="35537"),
                          vectors=c("v_CA21_5817","v_CA21_599","v_CA21_8","v_CA21_4917"),
                          level='DA', use_cache = FALSE, geo_format = 'sf', quiet = TRUE)
```

```{r}
centroids <- st_centroid(census_data$geometry)
coordinates <- st_coordinates(centroids) |> data.frame() #making it a data.frame()
```

